---
title: "Analyses"
author: "Vera Weisbecker and Cruise Speck"
date: "5 December 2018"
output: html_document
---



## load required packages and processed data



```{r }
library(ape)
library(geiger)
library(phytools)
library(caper)
library(car)
library(multcomp)
library(geomorph)
library(lmSupport)
library(ggplot2)
library(nlme)
library(dplyr)
library(RColorBrewer)
library(surface)


#the "utilities" file contains functions so this analysis page isn't too cluttered
source("../Analyses/Utilities.r")
load(file = "../Data/Processed_Data_for_Pgls.rda")
data_raw=read.csv("../Data/raw_data.csv", header=TRUE)

```


#First, some basic data on the quality of the raw dataset: How many percent of the dataset entries are there for each of the entry types?

```{r}


DataQuality=as.data.frame(sapply(data_raw[,2:28], function(x) sum((!is.na(x))/nrow(data_raw))*100))
colnames(DataQuality)=c ("% in raw dataset")
  
```

#Phylogenetic signal of tail length, head and body length, and body weight

```{r}
# Double-checking tip labels and data row names are in the same order

tree$tip.label==rownames(data)

#the actual tree lengths result in a singular matrix that can't be decomposed; ultrametricizing the tree helps
brlentree=compute.brlen(tree)
brlentree$tip.label==rownames(data)

phylosig(brlentree, data$Tl_mm, method = "K", nsim=1000)
phylosig(brlentree, data$Bl_mm, method="K" ,nsim=1000)
phylosig(brlentree, data$Wt_g, method="K", nsim=1000)


#lambda works on the tree but gives a different, very high phylogenetic signal

phylosig(tree, data$Tl_mm, method = "lambda", nsim=1000)
phylosig(tree, data$Bl_mm, method="lambda", nsim=1000)
phylosig(tree, data$Wt_g, method="lambda", nsim=1000)

```


# Testing the hypothesis that within-species body weight maxima are greater relative to body weight minima(i.e. the range is relatively wider), compared to the same ratio in within-species body or tail lengths) - for those species where ranges are available, body mass ranges are substantially greater

```{r}
#Marsupials

  ##return the min/max values to their un-logged form
  ratio_Tl=c(10^(dataFull$Tl_max)/10^(dataFull$Tl_min))
  ratio_Bl=c(10^(dataFull$Bl_max)/10^(dataFull$Bl_min))
  ratio_Wt=c(10^(dataFull$Wt_max)/10^(dataFull$Wt_min))
  
  ##Visualizing means of the ratios 
  boxplot(ratio_Tl, ratio_Bl,ratio_Wt, names=c("Tl ratio", "Bl ratio", "Wt ratio"))
  
  
  
  ##The VarDist.test code is in utilities. It runs a Kruskal-Wallis test and then a pairwise wilcoxon rank sum test on the ratios
  ratios=c(ratio_Tl, ratio_Bl, ratio_Wt)
  
  MinMaxtest(ratios,3)
  
  
  ##This is not mainly because of the extreme ranges in bandicoots!
  
  ratio_Tl_Band<-ratio_Tl[ - (grep("Macrotis_leucura", dataFull$Species):grep("Echymipera_rufescens", dataFull$Species))]
  ratio_Bl_Band<-ratio_Bl[ - (grep("Macrotis_leucura", dataFull$Species):grep("Echymipera_rufescens", dataFull$Species))]
  ratio_Wt_Band<-ratio_Wt[ - (grep("Macrotis_leucura", dataFull$Species):grep("Echymipera_rufescens", dataFull$Species))]
  
  ratiosBand=c(ratio_Tl_Band, ratio_Bl_Band, ratio_Wt_Band)
  MinMaxtest(ratiosBand,3)
  
  par(mfrow=c(1,2))
  boxplot(ratio_Tl, ratio_Bl,ratio_Wt, xlab="ratios with bandicoots")
  boxplot(ratio_Tl_Band, ratio_Bl_Band,ratio_Wt_Band, xlab="ratios without bandicoots")



```

#is there a difference in tail/body length or tail/weigth relationships between the sexes? This is only possible in a non phylogenetic context

```{r}
#Re-arrange datasets to put males, femals and averages under each other, turning into a data frame for column names
  
  Males<-cbind(dataFull$Tl_Av_M,dataFull$Bl_Av_M,dataFull$Wt_Av_M, rep(1,length(dataFull$Tl_Av_M)))
  Females<-cbind(dataFull$Tl_Av_F,dataFull$Bl_Av_F,dataFull$Wt_Av_F, rep(2,length(dataFull$Tl_Av_F)))
  Averages<-cbind(dataFull$Tl_mm, dataFull$Bl_mm,dataFull$Wt_g, rep(3,length(dataFull$Bl_mm)))
  
  MarsCombine=rbind(Males, Females,Averages)
  MarsCombine=as.data.frame(MarsCombine)
  colnames(MarsCombine) <-c("Tl","Bl","Wt","Sex")
  MarsCombine$Sex<-as.factor(MarsCombine$Sex)

  
#Are there differences in slope or intercept of Tail  and body length or weight according to sex, or averages between sexes? Using Anova from the car packages because this allows type III ANOVAs where the contribution of all other effect is adjusted for while computing each individual effect
  
  ## Body length:
  
    ###check model residuals before proceeding; some low-leverage outliers which are here accepted
    plot(lm(MarsCombine$Tl~MarsCombine$Bl*MarsCombine$Sex))
    
    ### Do regression slopes differ between sexes? 
    Anova(lm(MarsCombine$Tl~MarsCombine$Bl*MarsCombine$Sex),type = 3)
    
    
    ### No significant interactions, therefore dropping interactions to look for differences in intercept
    Anova(lm(MarsCombine$Tl~MarsCombine$Bl+MarsCombine$Sex),type = 3)
  
    
  ##Weight
  
    ###check model residuals before proceeding; no substantial outliers
    plot(lm(MarsCombine$Tl~MarsCombine$Wt*MarsCombine$Sex))
  
    ### No regression slope differences between sexes; no intercept differences either (so smaller females also have commensurately smaller tails it seems, regardless of weight or body length)
    Anova(lm(MarsCombine$Tl~MarsCombine$Wt*MarsCombine$Sex),type = 3)
    
    Anova(lm(MarsCombine$Tl~MarsCombine$Wt+MarsCombine$Sex),type = 3)
  
   
```

# How do tail lengths scale with body length and weight? 

```{r}

#Marsupials
    
      ##Tail vs. body length
      MarsTB=find.best.model(Tl_mm~Bl_mm,marsTree,mars);MarsTB$AIC_Weights
      #Checking appropriate distribution of model residuals      
      qqnorm (MarsTB$Models$Brownian)      
      MarsTB$Model_summaries$Brownian$tTable
      
      
      ##Tail vs body weight
      MarsTW=find.best.model(Tl_mm~Wt_g,marsTree,mars);MarsTW$AIC_Weights
      #Checking appropriate distribution of model residuals
      qqnorm (MarsTB$Models$Brownian)
      MarsTW$Model_summaries$Brownian$tTable
      
 ##rodents with Rattus
      
      ###Tail vs. body length
      rodTB=find.best.model(Tl_mm~Bl_mm,rodTree,rod);rodTB$AIC_Weights
      #Checking appropriate distribution of model residuals      
      qqnorm (rodTB$Models$Brownian)      
      rodTB$Model_summaries$Brownian$tTable
      
      
      ###Tail vs body weight
      rodTW=find.best.model(Tl_mm~Wt_g,rodTree,rod);rodTW$AIC_Weights
      #Checking appropriate distribution of model residuals
      qqnorm (rodTB$Models$Brownian)
      rodTW$Model_summaries$Brownian$tTable
      
  ##rodents without Rattus
      
      ###Tail vs. body length
      MurTB=find.best.model(Tl_mm~Bl_mm,murine_tree,murine_data);MurTB$AIC_Weights
      #Checking appropriate distribution of model residuals; excluding Rattus improves this     
      qqnorm (MurTB$Models$Brownian) 
      #but no change in scaling coefficient
      MurTB$Model_summaries$Brownian$tTable
      
      
      ###Tail vs body weight
      MurTW=find.best.model(Tl_mm~Wt_g,murine_tree,murine_data);MurTW$AIC_Weights
      #Checking appropriate distribution of model residuals
      qqnorm (MurTW$Models$Brownian_GrafScale)
      MurTW$Model_summaries$Brownian$tTable
      
```

# Does the significance level drop if body weight and body length are both part of the model?

```{r}
      
#Marsupials

      MarsTBW=find.best.model(Tl_mm~Bl_mm*Wt_g,marsTree,mars);MarsTW$AIC_Weights
      #Checking appropriate distribution of model residuals
      qqnorm (MarsTBW$Models$Brownian)
      anova(MarsTBW$Models$Brownian, type="marginal")
      
      ##dropping the interaction 
      MarsTBWNoInter=find.best.model(Tl_mm~Bl_mm+Wt_g,marsTree,mars);MarsTW$AIC_Weights
      #Checking appropriate distribution of model residuals
      qqnorm (MarsTBWNoInter$Models$Brownian)
      anova(MarsTBWNoInter$Models$Brownian, type="marginal")


```

# Aside from the association between body length and tail length,does locomotor mode play a role?

```{r}

#Marsupials
    
  ##Tail vs body length
    TB=find.best.model(Tl_mm~Bl_mm*Locomotor_use,marsTree,mars);TB$AIC_Weights
    TB= (gls(Tl_mm~Bl_mm*Locomotor_use, correlation=corBrownian(1,phy=marsTree), data=mars))    

       
    #check if the residuals of the model are OK, this is using qqnorm.gls from the nlme package
    qqnorm (TB$Models$Brownian)
    
    ###look at the model summary of the most likely model, e.g. 
    TB$Model_summaries$corBrownian$tTable
      
    ###if the interactions are not significant (i.e. slopes are not sig. different), drop them    
    TB_no_inter=find.best.model(Tl_mm~Bl_mm+Locomotor_use,marsTree,mars);TB$AIC_Weights
    TB_no_inter$Model_summaries$corBrownian$tTable
    anova(TB_no_inter$Model_summaries$corBrownian, type="marginal")
   
     #or use phytools - not sure how to deal with the label warnings
    Locomotor_use=c(mars$Locomotor_use)
    names(Locomotor_use)<-names(mars$Species)
    phylANOVA(marsTree,Locomotor_use,mars$Tl_mm)
    
     ###using the smaller, m/f specific datasets note THIS NEEDS LARGE NUMBERS OF DATA SO DOES NOT ALWAYS DO INTERACTIONS
    TBF=find.best.model(Tl_Av_F ~Bl_Av_F+Locomotor_use,mf_marstree, mf_mars);TB$AIC_Weights
    anova(TBF$Model_summaries$corPagel, type="marginal")
    
    TBM=find.best.model(Tl_Av_M~Bl_Av_M+Locomotor_use,mf_marstree,mf_mars); TB$AIC_Weights
    anova(TBM$Model_summaries$corPagel, type="marginal")
    
  ##Tail vs body WEIGHT
    TW=find.best.model(Tl_mm~Wt_g*Locomotor_use,marsTree,mars);TW$AIC_Weights
    anova(TW$Model_summaries$corBrownian, type="marginal")  
    
    ###no interactions, therefore
    TW_no_inter=find.best.model(Tl_mm~Wt_g+Locomotor_use,marsTree,mars);TW$AIC_Weights
    anova(TW_no_inter$Model_summaries$corBrownian, type="marginal") 
    
    
    ###using the smaller, m/f specific datasets
    TWF=find.best.model(Tl_Av_F ~Wt_Av_F+Locomotor_use,mf_marstree, mf_mars);TW$AIC_Weights
    anova(TWF$Model_summaries$corBrownian, type="marginal") 
      
    TWM=find.best.model(Tl_Av_M~Wt_Av_M+Locomotor_use,mf_marstree,mf_mars);TW$AIC_Weights
    anova(TWF$Model_summaries$corBrownian, type="marginal") 
  
    
#Rodents
    
   #All rodents 
    
    ##Tail vs body LENGTH
      RodTB=find.best.model(Tl_mm~Bl_mm*Locomotor_use,rodTree,rod);RodTB$AIC_Weights
      #several similarly likely models but they yield similar levels of significance
      qqnorm (RodTB$Models$Pagel_GrafScale)
      anova(RodTB$Models$Pagel)
      
      ##No interactions (i.e. no differences in tail/Bl slope according to locomotion), so dropping interaction
      RodTBNoInter=find.best.model(Tl_mm~Bl_mm+Locomotor_use,rodTree,rod);RodTBNoInter$AIC_Weights
      #qqnorm (RodTBNoInter$Models$Pagel_GrafScale)
      anova(RodTBNoInter$Models$Pagel_GrafScale)
        
     
    ##Tail vs body WEIGHT
      RodTW=find.best.model(Tl_mm~Wt_g*Locomotor_use,rodTree,rod);RodTW$AIC_Weights
      qqnorm (RodTWNoInter$Models$Pagel_GrafScale)
      anova(RodTW$Models$Pagel_GrafScale) 
      
      ##No interactions (i.e. no differences in tail/Bl slope according to locomotion), so dropping interaction
      RodTWNoInter=find.best.model(Tl_mm~Wt_g*Locomotor_use,rodTree,rod);RodTWNoInter$AIC_Weights
      anova(RodTBNoInter$Models$Pagel_GrafScale)
  
  #All rodents 
    
    ##Tail vs body LENGTH
      RodTB=find.best.model(Tl_mm~Bl_mm*Locomotor_use,rodTree,rod);RodTB$AIC_Weights
      #several similarly likely models; technically the Pagel correlation structure without Grafen-scaled trees performs slightly better but the level of significance is the same in just Pagel and the original tree (which is also the more likely model in the no interaction model)
      anova(RodTB$Models$Pagel_GrafScale)
      
      ##No interactions (i.e. no differences in tail/Bl slope according to locomotion), so dropping interaction
      RodTBNoInter=find.best.model(Tl_mm~Bl_mm+Locomotor_use,rodTree,rod);RodTBNoInter$AIC_Weights
      anova(RodTBNoInter$Models$Pagel_GrafScale)
        
     
    ##Tail vs body WEIGHT
      RodTW=find.best.model(Tl_mm~Wt_g*Locomotor_use,rodTree,rod);RodTW$AIC_Weights
      anova(RodTW$Models$Pagel_GrafScale) 
      
      ##No interactions (i.e. no differences in tail/Bl slope according to locomotion), so dropping interaction
      RodTWNoInter=find.best.model(Tl_mm~Wt_g*Locomotor_use,rodTree,rod);RodTWNoInter$AIC_Weights
      anova(RodTBNoInter$Models$Pagel_GrafScale)   
      
   
```


#Plotting male, female and averages 

```{r}
#Using the SURFACE package to look for significant shifts in evolutoinary regime
  
ForSurf=as.data.frame(cbind(mars$Tl_mm, mars$Bl_mm))
colnames(ForSurf) <- c("T1", "BL")
rownames(ForSurf) <- rownames(mars)

MarsSurfTree<-nameNodes(marsTree)
olist<-convertTreeData(MarsSurfTree,ForSurf)
otree<-olist[[1]]; odata<-olist[[2]]

fwd<-surfaceForward(otree, odata)
k<-length(fwd)

bwd<-surfaceBackward(otree, odata, starting_model = fwd[[k]], aic_threshold = 0, only_best = TRUE, verbose = FALSE, plotaic = FALSE)
bsum<-surfaceSummary(bwd)
kk<-length(bwd)

fsum=surfaceSummary(fwd)
bsum <- surfaceSummary(bwd)

surfaceTreePlot(MarsSurfTree, bwd[[kk]], labelshifts = T)

```








```{r}
  
#plot male-only, female-only, and averaged datasets together - Marsupials only 

par(mfrow=c(1,2))

plot(mars$Bl_Av_M,mars$Tl_Av_M, col="red", pch=19);points(mars$Bl_Av_F,mars$Tl_Av_F, col="blue", pch=19);points(mars$Bl_mm,mars$Tl_mm, col="green", pch=19);abline(lm(mars$Tl_Av_M~mars$Bl_Av_M), col="red");abline(lm(mars$Tl_Av_F~mars$Bl_Av_F), col="blue"); abline(lm(mars$Tl_mm~mars$Bl_mm), col="green")


plot(mars$Wt_Av_M,mars$Tl_Av_M, col="red", pch=19);points(mars$Wt_Av_F,mars$Tl_Av_F, col="blue", pch=19);points(mars$Wt_g,mars$Tl_mm, col="green", pch=19);abline(lm(mars$Tl_Av_M~mars$Wt_Av_M), col="red");abline(lm(mars$Tl_Av_F~mars$Wt_Av_F), col="blue"); abline(lm(mars$Tl_mm~mars$Wt_g), col="green")  


```

#Plotting according to tail use and phylogeny - body length

```{r}
#Marsupials
  
  ## initial setup: colours and margins
    colsOrderM=brewer.pal(n=length(levels(mars$Order)), name="Dark2")
    colsLoco=brewer.pal(n=length(levels(mars$Locomotor_use)), name="Dark2")
    colsOrderR=brewer.pal(n=length(levels(rodFull$Order)), name="Dark2")
    # Size of points
    cexes <- 1
    # this adjusts the titles of the double panel
    adjs <- -0.3
    # spaces to the left of text
    xinter <- 0.2
     # spaces between legend lines
    yInter <- 0.25
    #size of the box is determined by text width
    textwidth <- 0.2
    #insets the box a bit
    inset <- 0.01
  
    
    #4 panels, with margins of 2 lines
    par(mfrow=c(2,2), mar = c(4,4,4,4))
  
  plot(mars$Tl_mm~mars$Bl_mm, col=colsOrderM[mars$Order], pch=19, xlab="log Body length", ylab="log Tail length") ;legend("bottomright", c("Dasyuromorph" ,  "Diprotodontian", "Peramelemorph"),inset=inset, pch=19, col=colsOrderM, cex=cexes,y.intersp = yInter, x.intersp = xinter,text.width = textwidth+0.1, bty="n" ); #text(mars$Tl_mm~mars$Bl_mm, labels=mars$Species, cex=0.5)
  
  plot(mars$Tl_mm~mars$Bl_mm, col=colsLoco[mars$Locomotor_use], pch=19, xlab="log Body length", ylab="log Tail length" ); legend("bottomright",inset=inset,levels(mars$Locomotor_use) , pch=19, col=colsLoco, cex=cexes, x.intersp = xinter, y.intersp = yInter, text.width = textwidth, bty="n")
  
  mtext("Marsupials", adj=adjs,side=3, line=2, cex=1)
  
  ##Rodents
  
  plot(rodFull$Tl_mm~rodFull$Bl_mm, col=colsOrderR[rodFull$Order], pch=19, xlab="log Body length", ylab="log Tail length" );legend("bottomright", c("Murine" ,  "Rattus"),inset=inset, pch=19, col=colsOrderR, cex=cexes, x.intersp = xinter, y.intersp = yInter, text.width = textwidth-0.1, bty="n"); #text(rodFull$Tl_mm~rodFull$Bl_mm, labels=rodFull$Species, cex=cexes)
  
  plot(rodFull$Tl_mm~rodFull$Bl_mm, col=colsLoco[rodFull$Locomotor_use], pch=19, xlab="log Body length", ylab="log Tail length" ); legend("bottomright", inset=inset,levels(rodFull$Locomotor_use) , pch=19, col=colsLoco, cex=cexes,  ncol=2, x.intersp = xinter, y.intersp = yInter, text.width = textwidth-0.1, bty="n")
  
  mtext("Rodents", adj=adjs, line=2, cex = 1)

```


#Plotting according to tail use and phylogeny - body mass ADJUST ACCORDING TO ABOVE WHEN DONE

```{r}

  ##Marsupials
  
  
    ## initial setup: colours and margins
    colsOrderM=brewer.pal(n=length(levels(mars$Order)), name="Dark2")
    colsLoco=brewer.pal(n=length(levels(mars$Locomotor_use)), name="Dark2")
    colsOrderR=brewer.pal(n=length(levels(rodFull$Order)), name="Dark2")
    cexes<-0.6
    adjs=-0.3
    x.intersp=1
    
    #4 panels, with margins of 2 lines
    par(mfrow=c(2,2), mar = c(4,4,4,4))
  
  plot(mars$Tl_mm~mars$Wt_g, col=colsOrderM[mars$Order], pch=19, xlab="log Body mass", ylab="log Tail length" );legend("bottomright", c("Dasyuromorph" ,  "Diprotodontian", "Peramelemorph"), pch=19, col=colsOrderM, cex=cexes); #text(mars$Tl_mm~mars$Wt_g, labels=mars$Species, cex=0.5)
  
  plot(mars$Tl_mm~mars$Wt_g, col=colsLoco[mars$Locomotor_use], pch=19, xlab="log Body mass", ylab="log Tail mass" ); legend("bottomright",levels(mars$Locomotor_use) , pch=19, col=colsLoco, cex=cexes)
  
  mtext("Marsupials", adj=adjs,side=3, line=2, cex=1)
  
  ##Rodents
  
  plot(rodFull$Tl_mm~rodFull$Wt_g, col=colsOrderR[rodFull$Order], pch=19, xlab="log Body mass", ylab="log Tail length" );legend("bottomright", c("Murine" ,  "Rattus"), pch=19, col=colsOrderR, cex=cexes); #text(rodFull$Tl_mm~rodFull$Wt_g, labels=rodFull$Species, cex=cexes)
  
  plot(rodFull$Tl_mm~rodFull$Wt_g, col=colsLoco[rodFull$Locomotor_use], pch=19, xlab="log Body mass", ylab="log Tail length" ); legend("bottomright",levels(rodFull$Locomotor_use) , pch=19, col=colsLoco, cex=cexes)
  
  mtext("Rodents", adj=adjs, line=2, cex = 1)

```
