---
title: "Analyses"
author: "Vera Weisbecker and Cruise Speck"
date: "5 December 2018"
output: html_document
---



## load required packages and processed data



```{r }
library(ape)
library(geiger)
library(phytools)
library(caper)
library(car)
library(multcomp)
library(geomorph)
library(lmSupport)
library(ggplot2)
library(nlme)
library(dplyr)
library(RColorBrewer)
library(surface)
library(phia)


#the "utilities" file contains functions so this analysis page isn't too cluttered
source("../Analyses/Utilities.r")
load(file = "../Data/Processed_Data_for_Pgls.rda")
data_raw=read.csv("../Data/raw_data.csv", header=TRUE)

```


#First, some basic data on the quality of the raw dataset: How many percent of the dataset entries are there for each of the entry types?

```{r}


DataQuality=as.data.frame(sapply(data_raw[,2:28], function(x) sum((!is.na(x))/nrow(data_raw))*100))
colnames(DataQuality)=c ("% in raw dataset")
  
```

#Phylogenetic signal of tail length, head and body length, and body weight

```{r}
# Double-checking tip labels and data row names are in the same order

tree$tip.label==rownames(data)

#the actual tree lengths result in a singular matrix that can't be decomposed; ultrametricizing the tree helps
brlentree=compute.brlen(tree)
brlentree$tip.label==rownames(data)

phylosig(brlentree, data$Tl_mm, method = "K", nsim=1000)
phylosig(brlentree, data$Bl_mm, method="K" ,nsim=1000)
phylosig(brlentree, data$Wt_g, method="K", nsim=1000)


#lambda works on the tree but gives a different, very high phylogenetic signal

phylosig(tree, data$Tl_mm, method = "lambda", nsim=1000)
phylosig(tree, data$Bl_mm, method="lambda", nsim=1000)
phylosig(tree, data$Wt_g, method="lambda", nsim=1000)

```


# Testing the hypothesis that within-species body weight maxima are greater relative to body weight minima(i.e. the range is relatively wider), compared to the same ratio in within-species body or tail lengths) - for those species where ranges are available, body mass ranges are substantially greater

```{r}
#Marsupials

  ##return the min/max values to their un-logged form
  ratio_Tl=c(10^(dataFull$Tl_max)/10^(dataFull$Tl_min))
  ratio_Bl=c(10^(dataFull$Bl_max)/10^(dataFull$Bl_min))
  ratio_Wt=c(10^(dataFull$Wt_max)/10^(dataFull$Wt_min))
  
  ##Visualizing means of the ratios 
  boxplot(ratio_Tl, ratio_Bl,ratio_Wt, names=c("Tl ratio", "Bl ratio", "Wt ratio"))
  
  
  
  ##The VarDist.test code is in utilities. It runs a Kruskal-Wallis test and then a pairwise wilcoxon rank sum test on the ratios
  ratios=c(ratio_Tl, ratio_Bl, ratio_Wt)
  
  MinMaxtest(ratios,3)
  
  
  ##This is not mainly because of the extreme ranges in bandicoots!
  
  ratio_Tl_Band<-ratio_Tl[ - (grep("Macrotis_leucura", dataFull$Species):grep("Echymipera_rufescens", dataFull$Species))]
  ratio_Bl_Band<-ratio_Bl[ - (grep("Macrotis_leucura", dataFull$Species):grep("Echymipera_rufescens", dataFull$Species))]
  ratio_Wt_Band<-ratio_Wt[ - (grep("Macrotis_leucura", dataFull$Species):grep("Echymipera_rufescens", dataFull$Species))]
  
  ratiosBand=c(ratio_Tl_Band, ratio_Bl_Band, ratio_Wt_Band)
  MinMaxtest(ratiosBand,3)
  
  par(mfrow=c(1,2))
  boxplot(ratio_Tl, ratio_Bl,ratio_Wt, xlab="ratios with bandicoots")
  boxplot(ratio_Tl_Band, ratio_Bl_Band,ratio_Wt_Band, xlab="ratios without bandicoots")



```

#is there a difference in tail/body length or tail/weigth relationships between the sexes? This is only possible in a non phylogenetic context

```{r}
#Re-arrange datasets to put males, femals and averages under each other, turning into a data frame for column names
  
  Males<-cbind(dataFull$Tl_Av_M,dataFull$Bl_Av_M,dataFull$Wt_Av_M, rep(1,length(dataFull$Tl_Av_M)))
  Females<-cbind(dataFull$Tl_Av_F,dataFull$Bl_Av_F,dataFull$Wt_Av_F, rep(2,length(dataFull$Tl_Av_F)))
  Averages<-cbind(dataFull$Tl_mm, dataFull$Bl_mm,dataFull$Wt_g, rep(3,length(dataFull$Bl_mm)))
  
  MarsCombine=rbind(Males, Females,Averages)
  MarsCombine=as.data.frame(MarsCombine)
  colnames(MarsCombine) <-c("Tl","Bl","Wt","Sex")
  MarsCombine$Sex<-as.factor(MarsCombine$Sex)

  
#Are there differences in slope or intercept of Tail  and body length or weight according to sex, or averages between sexes? Using Anova from the car packages because this allows type III ANOVAs where the contribution of all other effect is adjusted for while computing each individual effect
  
  ## Body length:
  
    ###check model residuals before proceeding; some low-leverage outliers which are here accepted
    plot(lm(MarsCombine$Tl~MarsCombine$Bl*MarsCombine$Sex))
    
    ### Do regression slopes differ between sexes? 
    Anova(lm(MarsCombine$Tl~MarsCombine$Bl*MarsCombine$Sex),type = 3)
    
    
    ### No significant interactions, therefore dropping interactions to look for differences in intercept
    Anova(lm(MarsCombine$Tl~MarsCombine$Bl+MarsCombine$Sex),type = 3)
  
    
  ##Weight
  
    ###check model residuals before proceeding; no substantial outliers
    plot(lm(MarsCombine$Tl~MarsCombine$Wt*MarsCombine$Sex))
  
    ### No regression slope differences between sexes; no intercept differences either (so smaller females also have commensurately smaller tails it seems, regardless of weight or body length)
    Anova(lm(MarsCombine$Tl~MarsCombine$Wt*MarsCombine$Sex),type = 3)
    
    Anova(lm(MarsCombine$Tl~MarsCombine$Wt+MarsCombine$Sex),type = 3)
  
   
```

#Plotting according to tail use, locomotor mode, and phylogeny - weight and length are very similar, so I suspect one can be taken out in the end.

```{r}

  
# initial setup: colours and margins
  colsOrderM=brewer.pal(n=length(levels(data$Order)), name="Dark2")
  colsLoco=brewer.pal(n=length(levels(data$Locomotor_use)), name="Dark2")
  ##Size of points
  cexes <- 1
  ## this adjusts the titles of the double panel
  adjs <- -0.3
  ## spaces to the left of text
  xinter <- 0.2
  ## spaces between legend lines
  yInter <- 0.25
  ##size of the box is determined by text width
  textwidth <- 0.2
  ##insets the box a bit
  inset <- 0.01
  
    
    ##4 panels, with margins of 2 lines
    par(mfrow=c(2,2), mar = c(4,4,4,4))
  
plot(data$Tl_mm~data$Bl_mm, col=colsOrderM[data$Order], pch=19, xlab="log Body length", ylab="log Tail length") ;legend("bottomright", c("Dasyuromorph" ,  "Macropod", "Peramelemorph", "Possum"),inset=inset, pch=19, col=colsOrderM, cex=cexes,y.intersp = yInter, x.intersp = xinter,text.width = textwidth+0.1, bty="n" ); text(data$Tl_mm~data$Bl_mm, labels=data$Species, cex=0.5)
  
  plot(data$Tl_mm~data$Bl_mm, col=colsLoco[data$Locomotor_use], pch=19, xlab="log Body length", ylab="log Tail length" ); legend("bottomright",inset=inset,levels(data$Locomotor_use) , pch=19, col=colsLoco, cex=cexes, x.intersp = xinter, y.intersp = yInter, text.width = textwidth, bty="n");text(data$Tl_mm~data$Bl_mm, labels=data$Species, cex=0.5)
  
plot(data$Tl_mm~data$Wt_g, col=colsOrderM[data$Order], pch=19, xlab="log Weight (g)", ylab="log Tail length") ;legend("bottomright", c("Dasyuromorph" ,  "Diprotodontian", "Peramelemorph"),inset=inset, pch=19, col=colsOrderM, cex=cexes,y.intersp = yInter, x.intersp = xinter,text.width = textwidth+0.1, bty="n" ); #text(data$Tl_mm~data$Wt_g, labels=data$Species, cex=0.5)
  
  plot(data$Tl_mm~data$Wt_g, col=colsLoco[data$Locomotor_use], pch=19, xlab="log Weight (g)", ylab="log Tail length" ); legend("bottomright",inset=inset,levels(data$Locomotor_use) , pch=19, col=colsLoco, cex=cexes, x.intersp = xinter, y.intersp = yInter, text.width = textwidth, bty="n")
  


```



# How do tail lengths scale with body length and weight? Which predicts Tail length better? The find. best.model code finds the best fitting tree branch and phylogenetic correlation matrix for each model and allows to run an anova on just that.For details, see utilities.

```{r}
  
  ##Tail vs. body length: Significant and very high correlation coefficient (0.87)
  TB=find.best.model(Tl_mm~Bl_mm,tree,data);TB$AIC_Weights
  #Checking appropriate distribution of model residuals; , this is using qqnorm.gls from the nlme package. They mostly fall on a line    
  qqnorm (TB$Models$Brownian)      
  TB$Model_summaries$Brownian$tTable

  
  ##Tail vs body weight: significant but low correlation coefficient (0.33)
  TW=find.best.model(Tl_mm~Wt_g,tree,data);TW$AIC_Weights
  #Checking appropriate distribution of model residuals
  qqnorm (TB$Models$Brownian)
  TW$Model_summaries$Brownian$tTable
  
  #compute relative probabilities betweeen 0 and 100 by comparing AICS according to Burnham and Anderson
  AICs=c(TB$Model_summaries$Brownian$AIC,TW$Model_summaries$Brownian$AIC )
  AICmin=AICs-min(AICs)
  W=exp(-0.5*AICmin)/sum(exp(-0.5*AICmin))
  
  #Not surprisingly, body length is a far better predictor of body length:
  W 
  
```

# Weight remains a significant predictor of tail length even when correcting for body mass, but with a very low coefficient

```{r}
      
      TBW=find.best.model(Tl_mm~Bl_mm*Wt_g,tree,data);TBW$AIC_Weights
      #Checking appropriate distribution of model residuals
      qqnorm (TBW$Models$Brownian)
      anova(TBW$Models$Brownian, type="marginal")
      
      ##dropping the interaction 
      TBWNoInter=find.best.model(Tl_mm~Bl_mm+Wt_g,tree,data);TBWNoInter$AIC_Weights
      #Checking appropriate distribution of model residuals; some low-leverage outliers deemed acceptable in this context
      qqnorm (TBWNoInter$Models$Brownian)
      anova(TBWNoInter$Models$Brownian, type="marginal")
      
      TBWNoInter$Models$Brownian$coefficients

      
```

# Aside from the association between body length and tail length,does locomotor mode play a role? 


```{r}

#There are strong interactions between locomotor mode

TB=find.best.model(Tl_mm~Bl_mm*Locomotor_use,tree,data);TB$AIC_Weights

#check if the residuals of the model are OK; the fit is quite good
#qqnorm (TB$Models$Brownian)

###look at the model summary of the most likely model: there are significant interactions 
anova(TB$Models$Brownian)

#It is not possible to do post-hoc test in the pgls framework, so instead looking for interactions in the non-phylogenetically corrected lm analysis and using testInteractions from the phia package. Also refer to the plot. 

TB_lm <- lm(Tl_mm~Bl_mm*Locomotor_use, data=data)
testInteractions(TB_lm, adjustment="holm")
```
#This pattern is not driven by the anomalies in Peramelemorpha and the genus Macropus?
```{r}
which(data$Species=="Macropus_robustus")
which (data$Species=="Macropus_giganteus")

NoPerams_Roos=data[-c(which(data$Order =="Peramelemorph"), 25:36), ]

NoPerams_RoosTree=drop.tip(tree, name.check(tree,NoPerams_Roos)$tree_not_data )

#The tree contains dasyurids and peramelids, and non-pentapedal macropods but few possums
plot(NoPerams_RoosTree)

#Testing without interactions
TBNPR=find.best.model(Tl_mm~Bl_mm*Locomotor_use,NoPerams_RoosTree,NoPerams_Roos);TBNPR$AIC_Weights

#this is quite good and seems to fit the data better than the model with interaction
#qqnorm (TBNAPNoInter$Models$Brownian)

###Still a significant interaction between locomotor uses
anova(TBNPR$Models$Brownian, type="marginal")


```

#Arboreal species (mostly possums) scale differently to terrestrial species (mostly dasyurids and peramelids), but not scansorial (mostly dasyurid) or pentapedal (large macropod) species. This makes sense because pentapedal and arobreal species use their tail heavily in locomotion. This, and the fact that there are two instances each of arboreal and pentapedal locomotion in a significant interaction, suggests to us that it is these two that are differend from the rest.  However, it could be due to the strong phylogenetic signal of locomotion. Fitting a model that uses phylogenetic distinctions does not result in significant interactions between major clade (order or diprotodontian family):

```{r}

TB=find.best.model(Tl_mm~Bl_mm*Order,tree,data);TB$AIC_Weights

#check if the residuals of the model are OK; the fit is quite good
#qqnorm (TB$Models$Brownian)

###look at the model summary of the most likely model: there are significant interactions 
anova(TB$Models$Brownian)

#a post-hoc test of the uncorrected lm model also suggests no differences in scaling slope:

anova(TB_lm <- lm(Tl_mm~Bl_mm*Order, data=data))

```

#To look for differences in means, which may also contribute to differences in relative tail length between locomotor modes, we separate arboreal and pentapedal species (i.e. the those categories that had interaction terms with several other locomotor modes) to look at just saltators, scansorial, and terrestrial species. This shows that non-pentapedal, non-arboreal marsupials - even if including bounding but not pentapedal kangaroos - share the same scaling slope.

```{r}

#prepare tree
NonArbPent=data[c(which(data$Locomotor_use =="Terrestrial"), which ( data$Locomotor_use =="Scansorial"), which(data$Locomotor_use =="Saltatorial")),]

NonArbPentTree=drop.tip(tree, name.check(tree,NonArbPent)$tree_not_data )

#The tree contains dasyurids and peramelids, and non-pentapedal macropods but few possums
plot(NonArbPentTree)

#Testing without interactions
TBNAP=find.best.model(Tl_mm~Bl_mm+Locomotor_use,NonArbPentTree,NonArbPent);TBNAPNoInter$AIC_Weights

#this is quite good and seems to fit the data better than the model with interaction
#qqnorm (TBNAPNoInter$Models$Brownian)

###No differences in intercept!!
anova(TBNAP$Models$Brownian, type="marginal")

```
#SOMETHIGN WRONG, FEWER TIP LABELS THAN SPECIES FOR REASONS i DON'TBut again, is this due to the very low intercept Peramelemorphs having an undue effect on the slope of terrestrial species? It should not in the framework of pgls though.
```{r}

#prepare tree
NonArbPentPeram=data[c(which(data$Locomotor_use =="Terrestrial"), which ( data$Locomotor_use =="Scansorial"), which(data$Locomotor_use =="Saltatorial"), which(data$Order=="Peramelemorph")),]
#XXXXXXXXXX
NonArbPentPeramTree=drop.tip(tree, name.check(tree,NonArbPentPeram)$tree_not_data ))

NonArbPentPeramTree$tip.label==NonArbPentPeram$Species

#The tree contains dasyurids and peramelids, and non-pentapedal macropods but few possums
plot(NonArbPentPeramTree)

#MOdel
TBNAPP=find.best.model(Tl_mm~Bl_mm+Locomotor_use,NonArbPentPeramTree,NonArbPentPeram);TBNAPP$AIC_Weights

#this is quite good and seems to fit the data better than the model with interaction
#qqnorm (TBNAPNoInter$Models$Brownian)

###No differences in intercept!!
anova(TBNAP$Models$Brownian, type="marginal")

```

#Arboreal and pentapedal species also share their intercept

```{r}

#prepare tree
ArbPent=data[-c(which(data$Locomotor_use =="Terrestrial"), which ( data$Locomotor_use =="Scansorial"), which(data$Locomotor_use =="Saltatorial")),]

ArbPentTree=drop.tip(tree, name.check(tree,ArbPent)$tree_not_data )

#The tree contains mostly Diprotodontia
plot(ArbPentTree)

#The model 
TBAP=find.best.model(Tl_mm~Bl_mm+Locomotor_use,ArbPentTree,ArbPent);TBAP$AIC_Weights

#this is quite good and seems to fit the data better than the model with interaction
qqnorm (TBAP$Models$Grafen)

###No differences in intercept here, either.
anova(TBAP$Models$Grafen, type="marginal")

```



#With body weight, there are two instances of arboreal scaling being different to two others; so again, assuming that arboreal species are the ones that are different

```{r}
#There are strong interactions between locomotor mode

TW=find.best.model(Tl_mm~Wt_g*Locomotor_use,tree,data);TW$AIC_Weights

#check if the residuals of the model are OK; the fit is quite good
qqnorm (TW$Models$Brownian)

###look at the model summary of the most likely model: there are also significant interactions here
anova(TW$Models$Brownian, type="marginal")

#again, use non-phylogenetic interaction test; with weight, only arboreal species seem different though

TW_lm <- lm(Tl_mm~Wt_g*Locomotor_use, data=data)
testInteractions(TW_lm, adjustment="holm")

```

# Removing the arboreal species, which we assume are the ones driving the interaction differences, reveals differences in mean betweeen terrestrial species and others. 

```{r}
#prepare tree
NonArb_Weight=data[-c(which(data$Locomotor_use =="Arboreal")),]

NonArb_WeightTree=drop.tip(tree, name.check(tree,NonArb_Weight)$tree_not_data )

#The tree contains dasyurids and peramelids, and non-pentapedal macropods but few possums
plot(NonArb_WeightTree)

#Model
TWNA=find.best.model(Tl_mm~Wt_g+Locomotor_use,NonArb_WeightTree,NonArb_Weight);TBNAP$AIC_Weights

#qqnorm fit is OK, again with just a couple of "tails"
#qqnorm (TWNA$Models$Brownian)

###Differences in tail/weight scaling here
anova(TWNA$Models$Brownian, type = "marginal")


#again, no post-hoc tests can be run on pgls analyses, so using non-phylogenetic one here

NonArb_Tukey=aov(Tl_mm~Wt_g+Locomotor_use, data=NonArb_Weight)

TukeyHSD(NonArb_Tukey, "Locomotor_use")

#The above gives an odd error message that I have not seen before; but clearly it is not ignoring the contribution of the weight data to the model, which is reassuring:

NonArb_Tukey1=aov(Tl_mm~Locomotor_use, data=NonArb_Weight)

TukeyHSD(NonArb_Tukey1)

```


 

```{r}
#Using the SURFACE package to look for significant shifts in evolutionary regime - this currently does not work because the matrices become singular and can't be decomposed. On a smaller dataset I noticed that the regimes are estimated differently when regression residuals are used.
  
ForSurf=as.data.frame(cbind(data$Tl_mm, data$Bl_mm))
colnames(ForSurf) <- c("T1", "BL")
rownames(ForSurf) <- rownames(data)

MarsSurfTree<-nameNodes(tree)
olist<-convertTreeData(MarsSurfTree,ForSurf)
otree<-olist[[1]]; odata<-olist[[2]]

fwd<-surfaceForward(otree, odata)
k<-length(fwd)

bwd<-surfaceBackward(otree, odata, starting_model = fwd[[k]], aic_threshold = 0, only_best = TRUE, verbose = FALSE, plotaic = FALSE)
bsum<-surfaceSummary(bwd)
kk<-length(bwd)

fsum=surfaceSummary(fwd)
bsum <- surfaceSummary(bwd)

surfaceTreePlot(MarsSurfTree, bwd[[kk]], labelshifts = T)

```








```{r}
  
#plot male-only, female-only, and averaged datasets together - Marsupials only 

par(mfrow=c(1,2))

plot(data$Bl_Av_M,data$Tl_Av_M, col="red", pch=19);points(data$Bl_Av_F,data$Tl_Av_F, col="blue", pch=19);points(data$Bl_mm,data$Tl_mm, col="green", pch=19);abline(lm(data$Tl_Av_M~data$Bl_Av_M), col="red");abline(lm(data$Tl_Av_F~data$Bl_Av_F), col="blue"); abline(lm(data$Tl_mm~data$Bl_mm), col="green")


plot(data$Wt_Av_M,data$Tl_Av_M, col="red", pch=19);points(data$Wt_Av_F,data$Tl_Av_F, col="blue", pch=19);points(data$Wt_g,data$Tl_mm, col="green", pch=19);abline(lm(data$Tl_Av_M~data$Wt_Av_M), col="red");abline(lm(data$Tl_Av_F~data$Wt_Av_F), col="blue"); abline(lm(data$Tl_mm~data$Wt_g), col="green")  


```




#Plotting according to tail use and phylogeny - body mass ADJUST ACCORDING TO ABOVE WHEN DONE

```{r}

 # initial setup: colours and margins
  colsOrderM=brewer.pal(n=length(levels(data$Order)), name="Dark2")
  colsLoco=brewer.pal(n=length(levels(data$Locomotor_use)), name="Dark2")
  ##Size of points
  cexes <- 1
  ## this adjusts the titles of the double panel
  adjs <- -0.3
  ## spaces to the left of text
  xinter <- 0.2
  ## spaces between legend lines
  yInter <- 0.25
  ##size of the box is determined by text width
  textwidth <- 0.2
  ##insets the box a bit
  inset <- 0.01
  
    
   

```

#some code for possibly later

```{r}

#or use phytools - not sure how to deal with the label warnings
Locomotor_use=c(data$Locomotor_use)
names(Locomotor_use)<-names(data$Species)
phylANOVA(tree,Locomotor_use,data$Tl_mm)

```
